Pangenome graph layout by Path-Guided Stochastic Gradient Descent

### 2023-07-28 manuscript/main.tex
clean

	Originating from empirical inspection of word frequency tables, Zipf's law states that a word with rank $n$ occurs $1/n$ times as the most frequent one.
    This law is modeled by the Zipf distribution, whose rank-frequency distribution is related to the inverse power law. % repetitive
	Sampling $s_j$ from a Zipf distribution fixed in the $s_i$'s path position space increases the possibility to draw a nucleotide position close to $s_i$.
	So there is a high chance to use small nucleotide distances $nd_{ij}$ to refine the layout of nodes comprising a few base pairs.
    The Zipf distribution is also long tailed, with many occurrences of low frequency events, which ensures that longer-range structures are represented the output embedding space.

@@

	Graph layout initialization can significantly influence the quality of the final layout.
	In the 1D implementation, nodes are placed in the same order as they appear in the input graph by default, although support for random layout initialization is also provided.
	%Random placement ensures that each run of the PG-SGD method could potentially yield a different result, thereby increasing the likelihood of obtaining a meaningful layout. @Andrea: What is it that you wanted to say here?
	In 2D, we offer several layout initialization techniques.
	One option is to position nodes in the first layout dimension according to their order in the input graph and to add either uniform or Gaussian noise in the second dimension.
	Another possibility is to arrange nodes along a Hilbert curve, which often favors the creation of planar final layouts.
    We additionally support fixing a set of node coordinates, such as those in a reference genome, in their order and position, which allows us to build reference-focused views of complex pangenomes. \FIXME{Is this only for 1D? It would be \textit{amazing} if it were possible to do in 2D, and it should be easy. -EG}

>>>>

	Originating from empirical inspection of word frequency tables, Zipf's law states that a word with rank $n$ occurs $1/n$ times as the most frequent one.
    This law is modeled by the Zipf distribution, whose rank-frequency distribution is related to the inverse power law. % repetitive
	Sampling $s_j$ from a Zipf distribution fixed in the $s_i$'s path position space increases the possibility to draw a nucleotide position close to $s_i$.
	So there is a high chance to use small nucleotide distances $nd_{ij}$ to refine the layout of nodes comprising a few base pairs.
    The Zipf distribution is also long tailed, with many occurrences of low frequency events, which ensures that longer-range structures are represented in the output embedding space.

@@

	Graph layout initialization can significantly influence the quality of the final layout.
	In the 1D implementation, nodes are placed in the same order as they appear in the input graph by default, although support for random layout initialization is also provided.
	In 2D, we offer several layout initialization techniques.
	One option is to position nodes in the first layout dimension according to their order in the input graph and to add either uniform or Gaussian noise in the second dimension.
	Another possibility is to arrange nodes along a Hilbert curve, which often favors the creation of planar final layouts.
    We additionally support fixing a set of node coordinates, such as those in a reference genome, in their order and position, which allows us to build reference-focused views of complex pangenomes. \FIXME{Is this only for 1D? It would be \textit{amazing} if it were possible to do in 2D, and it should be easy. -EG}
